---
title: "KCH-IE Stats and Visualisation"
author: "Phoebe Cheong"
date: "05/08/2022"
output: html_document
---
Files can be downloaded from:https://github.com/chikusquish/MHDprojectIE2022

## KCH-IE 1. Statistical tests

**Install packages** for this session. Packages are as shown in: https://github.com/chikusquish/MHDprojectIE2022/blob/main/R-pseudocodes

**Load libraries** for this session

```{r, results='hide', warning=FALSE, message=FALSE}
library(writexl)
library(readxl)
library(tidyr)
library(dplyr)
library(tibble)
library(ggplot2)
library(stringr)
library(tidyverse)
library(data.table)
library(ggpubr)
library(ggsci)
library(plyr)
library(rcompanion)
library(survminer)
library(survival)
library(stats)
load("KCH-IE_statsandfigures.RData")
```

Check normality of age 
```{r, warning=FALSE, message=FALSE}
kch$age<-as.numeric(kch$age)
shapiro.test(kch$age)
```

Association between age group and blood culture results

```{r, echo=FALSE}
chisq.test(kch$age_group, kch$blood_culture)
```
modify values for linear regression
```{r, warning=FALSE, message=FALSE}
kch$staphylococcus[kch$staphylococcus=="Staphylococcus"]<-"1"
kch$staphylococcus[is.na(kch$staphylococcus)]<-"0"
kch$streptococcus[kch$streptococcus=="Streptococcus"]<-"1"
kch$streptococcus[is.na(kch$streptococcus)]<-"0"
kch$enterococcus[kch1$enterococcus=="Enterococcus"]<-"1"
kch$enterococcus[is.na(kch$enterococcus)]<-"0"
kch$other_organism[kch$other_organism=="Other organism"]<-"1"
kch$other_organism[is.na(kch$other_organism)]<-"0"
kch$no_organism[kch$no_organism=="No organism"]<-"1"
kch$no_organism[is.na(kch$no_organism)]<-"0"
kch$staphylococcus<-as.numeric(kch$staphylococcus)
kch$streptococcus<-as.numeric(kch$streptococcus)
kch$enterococcus<-as.numeric(kch$enterococcus)
kch$other_organism<-as.numeric(kch$other_organism)
kch$no_organism<-as.numeric(kch$no_organism)
```
Use logistics regression to find out which age group is most associated with which type of positive blood culture results

```{r, warning=FALSE, message=FALSE}
staph_age<- glm(staphylococcus ~ age_group, data=kch,  family = binomial)
staph_age
summary(staph_age)
exp(coef(staph_age))

strep_age<- glm(streptococcus ~ age_group, data=kch,  family = binomial)
strep_age
summary(strep_age)
exp(coef(strep_age))

entero_age<- glm(enterococcus ~ age_group, data=kch,  family = binomial)
entero_age
summary(entero_age)
exp(coef(entero_age))
```

Association between prosthesis type and blood culture results

```{r, echo=FALSE}
chisq.test(kch$prosthesis_type, kch$blood_culture)
```

Use logistics regression to find out which prosthesis type is most associated with which type of positive blood culture results

```{r, warning=FALSE, message=FALSE}
staph_pros<- glm(staphylococcus ~ prosthesis_type, data=kch,  family = binomial)
staph_pros
summary(staph_pros)
exp(coef(staph_pros))

strep_pros<- glm(streptococcus ~ prosthesis_type, data=kch,  family = binomial)
strep_pros
summary(strep_pros)
exp(coef(strep_pros))

entero_pros<- glm(enterococcus ~ prosthesis_type, data=kch,  family = binomial)
entero_pros
summary(entero_pros)
exp(coef(entero_pros))
```

Association between IVDU and blood culture results

```{r, echo=FALSE}
chisq.test(kch$IVDU, kch$blood_culture)
```

Use logistics regression to find out which prosthesis type is most associated with which type of positive blood culture results

```{r, warning=FALSE, message=FALSE}
staph_ivdu<- glm(staphylococcus ~ IVDU, data=kch,  family = binomial)
staph_ivdu
summary(staph_ivdu)
exp(coef(staph_ivdu))

strep_ivdu<- glm(streptococcus ~ IVDU, data=kch,  family = binomial)
strep_ivdu
summary(strep_ivdu)
exp(coef(strep_ivdu))

entero_ivdu<- glm(enterococcus ~ IVDU, data=kch,  family = binomial)
entero_ivdu
summary(entero_ivdu)
exp(coef(entero_ivdu))
```

## 2. Data visualisation
Figure 3.1
```{r, warning=FALSE, message=FALSE}
bloodculturecases<-ftable(kch$blood_culture, kch$diagnosis)%>%as.data.frame

bloodculturecases$blood_culture<-bloodculturecases$Var1
bloodculturecases$diagnosis<-bloodculturecases$Var2

cases<-ggplot(bloodculturecases, aes(x=blood_culture, y=Freq, fill=diagnosis))+
  geom_col(position="dodge",width=0.5,colour="black")+
  theme_minimal()+
  theme(axis.text.x = element_text(hjust = 0.5, size=11),
        axis.title=element_text(size=15), 
        legend.text = element_text(size=10))+
  scale_fill_manual(name="Diagnosis",values=c("yellow", "skyblue"), 
                    limits=c("Definite Endocarditis", "Possible Endocarditis"),
                    labels=c("Definite IE", "Possible IE"))+
  scale_x_discrete(limits=c("Streptococcus", "Staphylococcus", "Enterococcus", "Other organism", "No organism"))+
  ggtitle("Blood Culture Results of KCH Patients")+
  xlab("Blood culture results")+
  ylab("Number of patients")

cases
```

Figure 3.2a 
```{r, warning=FALSE, message=FALSE}
prosthesisbc1<-ftable(kch1$prosthesis_type, kch1$blood_culture)%>%as.data.frame()

prosthesisbc1$numberofpatients<-prosthesisbc1$Freq
prosthesisbc1$prosthesis_type<-as.character(prosthesisbc1$Var1)
prosthesisbc1$blood_culture<-prosthesisbc1$Var2

prosthesisbc1$prosthesis_type = factor(prosthesisbc1$prosthesis_type, levels=c("ICD/PPM", "TAVI", "Surgical Valve"))


prosthesisbccases1<-ggplot(prosthesisbc1, aes(x=blood_culture, y=Freq, fill=prosthesis_type))+
  geom_col(position="dodge",width=0.5,colour="black")+
  theme_minimal()+
  theme(axis.text.x = element_text(hjust = 0.5, size=11),
        axis.text.y = element_text(size=11),
        axis.title=element_text(size=15), 
        legend.text = element_text(size=10))+
  scale_fill_manual(name="Prosthesis Type", values=c("yellow", "skyblue", "purple"))+
  ggtitle("Blood Culture Results of Definite IE Patients by Prosthesis Type")+
  xlab("Blood Culture")+
  ylab("Number of patients")

prosthesisbccases1
```


Figure 3.2b
```{r, warning=FALSE, message=FALSE}
prosthesisbc2<-ftable(kch2$prosthesis_type, kch2$blood_culture)%>%as.data.frame()

prosthesisbc2$numberofpatients<-prosthesisbc2$Freq
prosthesisbc2$prosthesis_type<-as.character(prosthesisbc2$Var1)
prosthesisbc2$blood_culture<-prosthesisbc2$Var2

prosthesisbc2$prosthesis_type = factor(prosthesisbc2$prosthesis_type, levels=c("ICD/PPM", "TAVI", "Surgical Valve"))


prosthesisbccases2<-ggplot(prosthesisbc2, aes(x=blood_culture, y=Freq, fill=prosthesis_type))+
  geom_col(position="dodge",width=0.5,colour="black")+
  theme_minimal()+
  theme(axis.text.x = element_text(hjust = 0.5, size=11),
        axis.text.y = element_text(size=11),
        axis.title=element_text(size=15), 
        legend.text = element_text(size=10))+
  scale_fill_manual(name="Prosthesis Type", values=c("yellow", "skyblue", "purple"))+
  ggtitle("Blood Culture Results of Possible IE Patients by Prosthesis Type")+
  xlab("Blood Culture")+
  ylab("Number of patients")

prosthesisbccases2
```

Figure 3.3a
```{r, warning=FALSE, message=FALSE}
ivdubc1<-ftable(kch1$IVDU, kch1$blood_culture)%>%as.data.frame()

ivdubc1$numberofpatients<-ivdubc1$Freq
ivdubc1$IVDU<-as.character(ivdubc1$Var1)
ivdubc1$blood_culture<-ivdubc1$Var2

ivdubccases1<-ggplot(ivdubc1, aes(x=blood_culture, y=Freq, fill=IVDU))+
  geom_col(position="dodge",colour="black")+
  theme_minimal()+
  theme(axis.text.x = element_text(hjust = 0.5, size=11),
        axis.text.y = element_text(size=11),
        axis.title=element_text(size=15), 
        legend.text = element_text(size=10))+
  scale_fill_manual(name="Prosthesis Type", values=c("yellow", "skyblue", "purple"))+
  ggtitle("Blood culture results of Intravenous Drug Use in Definite IE patients")+
  xlab("Blood culture results")+
  ylab("Number of patients")

ivdubccases1
```

Figure 3.3b
```{r, warning=FALSE, message=FALSE}
ivdubc2<-ftable(kch2$IVDU, kch2$blood_culture)%>%as.data.frame()

ivdubc2$numberofpatients<-ivdubc2$Freq
ivdubc2$IVDU<-as.character(ivdubc2$Var1)
ivdubc2$blood_culture<-ivdubc2$Var2

ivdubccases2<-ggplot(ivdubc2, aes(x=blood_culture, y=Freq, fill=IVDU))+
  geom_col(position="dodge",colour="black")+
  theme_minimal()+
  theme(axis.text.x = element_text(hjust = 0.5, size=11),
        axis.text.y = element_text(size=11),
        axis.title=element_text(size=15), 
        legend.text = element_text(size=10))+
  scale_fill_manual(name="Prosthesis Type", values=c("yellow", "skyblue", "purple"))+
  ggtitle("Blood culture results of Intravenous Drug Use in Possible IE patients")+
  xlab("Blood culture results")+
  ylab("Number of patients")

ivdubccases2
```

Figure 3.4a
```{r, warning=FALSE, message=FALSE}
culpritpathogencases<-ftable(kch$culprit_pathogen, kch$`Probable origin (based off google)`)%>%as.data.frame
culpritpathogencases$Freq[culpritpathogencases$Freq=="0"]<-NA
culpritpathogencases<-na.omit(culpritpathogencases)
culpritpathogencases$culpritpathogen<-culpritpathogencases$Var1
culpritpathogencases$origin<-culpritpathogencases$Var2

cpcases<-ggplot(culpritpathogencases, aes(x=culpritpathogen, y=Freq, fill=origin))+
  geom_col(position="dodge",width=0.5,colour="black")+
  theme_minimal()+
  theme(axis.text.x = element_text(hjust=1.0, size=9, angle=90),
        axis.title=element_text(size=10), 
        legend.position = "top")+
  scale_fill_manual(values=c("orange", "violet", "brown", "grey", "blue", "skyblue", "purple", "white", "yellow", "deeppink","pink"))+
  ggtitle("Culprit Pathogens in KCH IE Patients")+
  xlab("Culprit Pathogen")+
  ylab("Number of patients")

cpcases
```

Figure 3.4b

```{r, warning=FALSE, message=FALSE}
origin1<-ftable(kch$`Probable origin (based off google)`, kch$diagnosis)%>%as.data.frame
origin1$Freq[origin1$Freq=="0"]<-NA
origin1<-na.omit(origin1)
origin1$origin<-origin1$Var1
origin1$diagnosis<-origin1$Var2

origincases1<-ggplot(origin1, aes(x=origin, y=Freq, fill=diagnosis))+
  geom_col(position="dodge",width=0.5,colour="black")+
  theme_minimal()+
  theme(axis.text.x = element_text(hjust=1.0, size=9, angle=90),
        axis.title=element_text(size=12), 
        legend.position = "top")+
  scale_fill_manual(name="Diagnosis",values=c("yellow", "skyblue"))+
  ggtitle("Origin of Culprit Pathogens in KCH Patients")+
  xlab("Origin")+
  ylab("Number of patients")

origincases1
```

Figure 3.5a
```{r, warning=FALSE, message=FALSE}
prosthesisgenus1<-ftable(kch1$prosthesis_type, kch1$genus)%>%as.data.frame()
prosthesisgenus1$Freq[prosthesisgenus1$Freq=="0"]<-NA
prosthesisgenus1<-na.omit(prosthesisgenus1)
prosthesisgenus1$numberofpatients<-prosthesisgenus1$Freq
prosthesisgenus1$prosthesis_type<-as.character(prosthesisgenus1$Var1)
prosthesisgenus1$genus<-prosthesisgenus1$Var2

prosthesisgenus1$prosthesis_type = factor(prosthesisgenus1$prosthesis_type, levels=c("ICD/PPM", "TAVI", "Surgical Valve"))

prosthesisgenuscases1<-ggplot(prosthesisgenus1, aes(x=genus, y=Freq, fill=prosthesis_type, drop=T))+
  geom_col(position="dodge",width=0.5,colour="black")+
  theme_minimal()+
  theme(axis.text.x = element_text(hjust = 0.5, size=11, angle=90),
        axis.text.y = element_text(size=11),
        axis.title=element_text(size=15), 
        legend.text = element_text(size=10))+
  scale_fill_manual(name="Prosthesis Type", values=c("yellow", "skyblue", "purple"))+
  ggtitle("Genus of Culprit Pathogens in Definite IE Patients by Prosthesis Type")+
  xlab("Genus")+
  ylab("Number of patients")

prosthesisgenuscases1
```

Figure 3.5b
```{r, warning=FALSE, message=FALSE}
prosthesisgenus2<-ftable(kch2$prosthesis_type, kch2$genus)%>%as.data.frame()
prosthesisgenus2$Freq[prosthesisgenus2$Freq=="0"]<-NA
prosthesisgenus2<-na.omit(prosthesisgenus2)
prosthesisgenus2$numberofpatients<-prosthesisgenus2$Freq
prosthesisgenus2$prosthesis_type<-as.character(prosthesisgenus2$Var1)
prosthesisgenus2$genus<-prosthesisgenus2$Var2

prosthesisgenus2$prosthesis_type = factor(prosthesisgenus2$prosthesis_type, levels=c("ICD/PPM", "TAVI", "Surgical Valve"))

prosthesisgenuscases2<-ggplot(prosthesisgenus2, aes(x=genus, y=Freq, fill=prosthesis_type, drop=T))+
  geom_col(position="dodge",width=0.5,colour="black")+
  theme_minimal()+
  theme(axis.text.x = element_text(hjust = 0.5, size=11, angle=90),
        axis.text.y = element_text(size=11),
        axis.title=element_text(size=15), 
        legend.text = element_text(size=10))+
  scale_fill_manual(name="Prosthesis Type", values=c("yellow", "skyblue", "purple"))+
  ggtitle("Genus of Culprit Pathogens in Possible IE Patients by Prosthesis Type")+
  xlab("Genus")+
  ylab("Number of patients")

prosthesisgenuscases2
```

## 3. Kaplan-Meier Survival Rate

Figure 3.6a
```{r, warning=FALSE, message=FALSE}
metadata$bloodculture<-factor(metadata$bloodculture, levels=c("No organism", "Other organism",  
                                                      "Staphylococcus", "Streptococcus", "Enterococcus"))

s365_all <- Surv(metadata$Time365, metadata$Death365)
survfit(s365_all~1)
survfit(Surv(Time365, Death365)~1, data=metadata)
sfit365_all <- survfit(Surv(Time365, Death365)~1, data=metadata)
sfit365_all <- survfit(Surv(Time365, Death365)~bloodculture, data=metadata)

ggsurvplot(sfit365_all, legend="right", xlim=c(0, 370))

survbc_all<-ggsurvplot(sfit365_all, pval=TRUE,xlim=c(0, 370), legend="right",
                      title="Kaplan-Meier Curve for Infective Endocarditis Survival \nin KCH endocarditis patients")
survbc_all
fit365_all <- coxph(Surv(Time365, Death365)~bloodculture, data=metadata)
fit365_all
summary(fit365_all)
```

Figure 3.6b
```{r, warning=FALSE, message=FALSE}
kch1$bloodculture<-factor(kch1b$bloodculture, levels=c("No organism", "Other organism",  
                                                      "Staphylococcus", "Streptococcus", "Enterococcus"))
s365_defIE <- Surv(kch1b$Time365, kch1b$Death365)
survfit(s365_defIE~1)
survfit(Surv(Time365, Death365)~1, data=kch1b)
sfit365_defIE <- survfit(Surv(Time365, Death365)~1, data=kch1b)
kch1b$dg<-kch1b$`Diagnosis Confirmation`
sfit365_defIE <- survfit(Surv(Time365, Death365)~bloodculture, data=kch1b)

ggsurvplot(sfit365_defIE, legend.position="right")

survbc_defIE<-ggsurvplot(sfit365_defIE, pval=TRUE,xlim=c(0, 370), legend="right",
           title="Kaplan-Meier Curve for Infective Endocarditis Survival \nin Definite Endocarditis patients")
survbc_defIE

fit365_defIE <- coxph(Surv(Time365, Death365)~bloodculture, data=kch1b)
fit365_defIE 
summary(fit365_defIE)
```

Figure 3.6c
```{r, warning=FALSE, message=FALSE}
kch2b$bloodculture<-factor(kch2b$bloodculture, levels=c("No organism", "Other organism", 
                                                      "Staphylococcus", "Streptococcus", "Enterococcus"))

s365_possIE <- Surv(kch2b$Time365, kch2b$Death365)
survfit(s365_possIE~1)
survfit(Surv(Time365, Death365)~1, data=kch2b)
sfit365_possIE <- survfit(Surv(Time365, Death365)~1, data=kch2b)

sfit365_possIE <- survfit(Surv(Time365, Death365)~bloodculture, data=kch2b)

ggsurvplot(sfit365_possIE, legend.position="right")

survbc_possIE<-ggsurvplot(sfit365_possIE, pval=TRUE,xlim=c(0, 370), legend="right",
                      title="Kaplan-Meier Curve for Infective Endocarditis Survival \nin Possible Endocarditis patients")
survbc_possIE
fit365_possIE <- coxph(Surv(Time365, Death365)~bloodculture, data=kch2b)
fit365_possIE 
summary(fit365_possIE)
```