---
title: "TMSC1TMHD 2021/2022 Dissertation Appendix File1 - STH Pseudocodes"
author: "Phoebe Cheong"
date: "04/08/2022"
output:
  html_document:
    toc: yes
    keep_md: yes
  word_document:
    toc: yes
---

Files can be downloaded from:https://github.com/chikusquish/MHDprojectIE2022

## STH-IE 1. Create phyloseq object

**Install packages** for this session. Packages are as shown in: https://github.com/chikusquish/MHDprojectIE2022/blob/main/R-pseudocodes

**Load libraries** for this session

```{r, results='hide', warning=FALSE, message=FALSE}
library(writexl)
library(readxl)
library(tidyr)
library(dplyr)
library(tibble)
library(ggplot2)
library(phyloseq)
library(microbiome)
library(vegan)
library(stringr)
library(tidyverse)
library(devtools)
library(pairwiseAdonis)
load("STH-IE create phyloseq object.RData")
```

The first step of creating the phyloseq object is to modify the merged_estimated_read.txt file.

```{r, results='hide', warning=FALSE, message=FALSE}
main=read.table('HHS_IE_merged_estimated_number_read.txt',header=TRUE,sep="\t")
t = separate(main,X.clade_name, into = c("Kingdom","Phylum", "Class", "Order", "Family", "Genus", "Species"), sep="\\|")
t=t[!is.na(t$Species), ]
t=t[!duplicated(t$Species),]
t = t %>% mutate(Kingdom = str_replace(Kingdom, 'k__',''),
                 Phylum = str_replace(Phylum, 'p__',''),
                 Class = str_replace(Class, 'c__',''),
                 Order = str_replace(Order, 'o__',''),
                 Family = str_replace(Family, 'f__',''),
                 Genus = str_replace(Genus, 'g__',''),
                 Species = str_replace(Species, 's__',''))
names(t)[names(t) == 'X272IE_S111_profile_IEcount.txt'] <- "272IE"
names(t)[names(t) == 'X273IE_S112_profile_IEcount.txt'] <- "273IE"
names(t)[names(t) == 'X290IE_S113_profile_IEcount.txt'] <- "290IE"
names(t)[names(t) == 'X293IE_S13_profile_IEcount.txt'] <- "293IE"
names(t)[names(t) == 'X294IE_S114_profile_IEcount.txt'] <- "294IE"
names(t)[names(t) == 'X295IE_S14_profile_IEcount.txt'] <- "295IE"
names(t)[names(t) == 'X296IE_S115_profile_IEcount.txt'] <- "296IE"
names(t)[names(t) == 'SRS013942_profile_count.txt'] <- "SRS013942"
names(t)[names(t) == 'SRS014468_profile_count.txt'] <- "SRS014468"
names(t)[names(t) == 'SRS014692_profile_count.txt'] <- "SRS014692"
names(t)[names(t) == 'SRS015055_profile_count.txt'] <- "SRS015055"
names(t)[names(t) == 'SRS019120_profile_count.txt'] <- "SRS019120"
names(t)[names(t) == 'SRS065518_profile_count.txt'] <- "SRS065518"
names(t)[names(t) == 'SRS104275_profile_count.txt'] <- "SRS104275"
names(t)[names(t) == 'SRS147126_profile_count.txt'] <- "SRS147126"

t=as.data.frame(t[,-which(names(t) %in% c("SRS015055", "SRS019120"))])
t = t %>% mutate(Genus = str_replace(Genus, 'Clostridiales_Family_XIII_Incertae_Sedis_unclassified','Clostridiales**'))

x=1:dim(t)[1]
OTUs= paste("OTU",x)
t=add_column(t, OTUs = OTUs, .before = "clade_taxid")
t[is.na(t)] = 0
write_xlsx(t,"~/Desktop/Mod5_Data_Analysis/HHS_IE_combined/HHS_IE_Estimated_number_TX.xlsx")
```

1.  Create CountMatrix

```{r, results='hide', warning=FALSE, message=FALSE}
OTU =as.data.frame(t[,-which(names(t) %in% c("Kingdom","Phylum", "Class", "Order", "Family", "Genus","Type","clade_taxid"))])
write_xlsx(OTU,"~/Desktop/Mod5_Data_Analysis/HHS_IE_combined/HHS_IE_OTU.xlsx")

row.names(OTU) = OTU$OTUs
OTU =as.data.frame(OTU[,-which(names(OTU) %in% c("OTUs","Species"))])
OTU [is.na(OTU)] = 0

names(OTU) = gsub(pattern = "_profile_IEcount.txt", replacement = "", x = names(OTU))
names(OTU) = gsub(pattern = "_profile_count.txt", replacement = "", x = names(OTU))

OTU_modify<-t(OTU)%>%as.data.frame
  
OTU_modify2<-OTU_modify[,-which(names(OTU_modify) %in% c("OTU 196","OTU 197", "OTU 198", "OTU 199", "OTU 200", "OTU 202", "OTU 205", 
                             "OTU 206", "OTU 207", "OTU 210", "OTU 212", "OTU 213", "OTU 214", "OTU 215",
                             "OTU 216", "OTU 219", "OTU 220"))]
OTU_modify3<-t(OTU_modify2)%>%as.data.frame
CountMatrix = OTU_modify3 %>% as.matrix()
mode(CountMatrix) <- 'integer'
```

2.  Create TaxaMatrix

```{r, results='hide', warning=FALSE, message=FALSE}
TAX =as.data.frame(t[,which(names(t) %in% c("Kingdom","Phylum", "Class", "Order", "Family", "Genus","Species","OTUs"))])
write_xlsx(TAX,"~/Desktop/Mod5_Data_Analysis/HHS_IE_combined/HHS_IE_TAXA.xlsx")

row.names(TAX) = TAX$OTUs
tax_modify = t(TAX)%>%as.data.frame
tax_modify2<-tax_modify[,-which(names(tax_modify) %in% c("OTU 196","OTU 197", "OTU 198", "OTU 199", "OTU 200", "OTU 202", "OTU 205", 
                                                         "OTU 206", "OTU 207", "OTU 210", "OTU 212", "OTU 213", "OTU 214", "OTU 215",
                                                         "OTU 216", "OTU 219", "OTU 220"))]
tax_modify3<-t(tax_modify2)%>%as.data.frame
tax_modify3 =as.data.frame(tax_modify3[,-which(names(tax_modify3) %in% c("OTUs"))])
TaxaMatrix <- tax_modify3 %>% as.matrix()
```

3.  Load metadata and create sampleData

```{r, results='hide', warning=FALSE, message=FALSE}
Metadata <- read_xlsx("~/Desktop/Mod5_Data_Analysis/HHS_IE_combined/Metadata_HHS_IE.xlsx", sheet="Merged_Projects") %>% as.data.frame()
rownames(Metadata) <- Metadata$SampleID
```

4.  Create phyloseq object from all three items above

```{r, results='hide', warning=FALSE, message=FALSE}
otuTABLE <- otu_table(CountMatrix, taxa_are_rows = TRUE)
taxTABLE <- tax_table(TaxaMatrix)
sampleDATA <- sample_data(Metadata)
```

```{r}
phylo_obj <- phyloseq(otuTABLE, taxTABLE, sampleDATA)
phylo_obj
```

## STH-IE 2. Run statistical tests on taxa results

Extract abundance profiles, metadata and microbiome taxonomy from phyloseq object
```{r, results='hide', warning=FALSE, message=FALSE}
OTUdata <- abundances(phylo_obj)#extract abundance profiles 
SampleData <- meta(phylo_obj)#extract meta data
TAXAData <- as.data.frame(tax_table(phylo_obj)@.Data)#extract taxonomy of microbiome

Adiv <- estimate_richness(phylo_obj,measures=c("Observed","Chao1","ACE","Shannon","Simpson","InvSimpson"))
rownames(Adiv)
row.names(Adiv) <- c("272IE", "273IE", "290IE", "293IE", "294IE", "295IE", "296IE", "SRS013942", "SRS014468", "SRS014692", "SRS065518", "SRS104275", "SRS147126")
Adiv$SampleID <- c("272IE", "273IE", "290IE", "293IE", "294IE", "295IE", "296IE", "SRS013942", "SRS014468", "SRS014692", "SRS065518", "SRS104275", "SRS147126")
write_xlsx(Adiv,"~/Desktop/Mod5_Data_Analysis/HHS_IE_combined/Adiv_HHS_IE.xlsx")
Adiv <- full_join(SampleData,Adiv,by="SampleID")#full_join function from dplyr
```

Insert Shannon Evenness Index

```{r, results='hide', warning=FALSE, message=FALSE}
H <- Adiv$Shannon
S1 <- Adiv$Observed
S <- log(S1)
evenness <- H/S
evenness
Adiv$Evenness = evenness
```

**Analyse alpha diversity**: Determine index for analysis

```{r}
index_defIE<-which(Adiv$IE_dx=="Definite_IE")
index_hrIE<-which(Adiv$IE_dx=="Highrisk_IE")
index_HHS<-which(Adiv$IE_dx=="Healthy")
```

Shapiro-Wilk test: check normality of data

```{r}
shapiro.test(Adiv$Shannon)#### p>0.05
shapiro.test(Adiv$Chao1)#### p>0.05
shapiro.test(Adiv$InvSimpson)#### p>0.05
shapiro.test(Adiv$Evenness)#### p>0.05
```

All alpha diversity measures are parametric, therefore ANOVA is used to analyse differences in mean.

```{r}
########## SHANNON ############
anova_adiv1<-aov(Adiv$Shannon ~ Adiv$IE_dx)
anova_adiv1
summary(anova_adiv1)
```

```{r}
tukey_IE_HHS1<-TukeyHSD(anova_adiv1)
tukey_IE_HHS1
```

```{r}
########### CHAO1 ############
anova_adiv2<-aov(Adiv$Chao1 ~ Adiv$IE_dx)
anova_adiv2
summary(anova_adiv2)
```

```{r}
tukey_IE_HHS2<-TukeyHSD(anova_adiv2)
tukey_IE_HHS2
```

```{r}
########### InvSimpson ############
anova_adiv3<-aov(Adiv$InvSimpson ~ Adiv$IE_dx)
anova_adiv3
summary(anova_adiv3)
```

```{r}
tukey_IE_HHS3<-TukeyHSD(anova_adiv3)
tukey_IE_HHS3
```

```{r}
########### Shannon Evenness Index ############
anova_adiv4<-aov(Adiv$Evenness ~ Adiv$IE_dx)
anova_adiv4
summary(anova_adiv4)
```

**Analyse beta diversity**

```{r, warning=FALSE, message=FALSE}
#log10 transformation for the abundance data
phyobj_shift<- microbiome::transform(phylo_obj,transform="log10")
#beta diversity based on euclidean distance 
Betdiv <- as.matrix(phyloseq::distance(phyobj_shift, method="bray"))
Betdiv[1:7, 1:7]
```

PERMANOVA analysis

```{r, warning=FALSE, message=FALSE}
SampleData2<-SampleData
set.seed(1)
Betdiv_HHS_IE<-adonis2(Betdiv ~ IE_dx, data=SampleData2, permutations=999)
Betdiv_HHS_IE
```

PERMANOVA posthoc

```{r, warning=FALSE, message=FALSE}
betdivposthoc_HHS_IE<-pairwise.adonis(Betdiv,factors=SampleData$IE_dx)
betdivposthoc_HHS_IE
```
