---
title: "STH-IE_visualisation"
author: "Phoebe Cheong"
date: "05/08/2022"
output: html_document
---
Files can be downloaded from:https://github.com/chikusquish/MHDprojectIE2022

## STH-IE 3. Create phyloseq object

**Install packages** for this session. Packages are as shown in: https://github.com/chikusquish/MHDprojectIE2022/blob/main/R-pseudocodes

**Load libraries** for this session

```{r, results='hide', warning=FALSE, message=FALSE}
library(phyloseq)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(ggsci)
library(stringr)
library(dutchmasters)
library(RColorBrewer)
load("STH-IE_visualisation.RData")
```

## 1. Create object for generation of figures

```{r}
phylo_obj@sam_data[1:3,1:3]

pd <- psmelt(phylo_obj)

```

Figure 3.7a

```{r, warning=FALSE, message=FALSE}
my_comparisons <- list( c("Definite_IE", "Healthy"), c("Definite_IE", "Highrisk_IE"), c("Healthy", "Highrisk_IE") )
boxplot_shannon<-ggplot(data = Adiv, aes(x = IE_dx, y = Shannon, fill=IE_dx)) + 
                        geom_boxplot(width=0.5)+
                        theme_minimal()+
                        theme(legend.position = "none")+
                        stat_summary(fun.y="mean", geom="point", shape=23, size=3, fill="white")+
                        geom_jitter(shape=1, position=position_jitter(0.1))+
                        ggtitle("Shannon Index by Diagnosis") +
                        scale_x_discrete(name= "Diagnosis", 
                                         limits=c("Definite_IE", "Healthy", "Highrisk_IE"),
                                         labels=c("Definite IE", "Healthy", "High risk IE"))+
                        scale_fill_manual(limits=c("Definite_IE", "Healthy", "Highrisk_IE"),
                                        values=c("skyblue","orange","purple"))+
                        stat_compare_means(comparisons=my_comparisons, label.y=c(4.2, 4.6, 4.4))

boxplot_shannon
```

Figure 3.7b

```{r, warning=FALSE, message=FALSE}
boxplot_chao1<- ggplot(data = Adiv, aes(x = IE_dx, y = Chao1, fill=IE_dx)) + 
                        geom_boxplot(width=0.5)+
                        theme_minimal()+
                        theme(legend.position = "none")+
                        stat_summary(fun.y="mean", geom="point", shape=23, size=3, fill="white")+
                        geom_jitter(shape=1, position=position_jitter(0.1))+
                        ggtitle("Chao1 Index by Diagnosis") +
                        scale_x_discrete(name= "Diagnosis", 
                                          limits=c("Definite_IE", "Healthy", "Highrisk_IE"),
                                          labels=c("Definite IE", "Healthy", "High risk IE"))+
                        scale_fill_manual(limits=c("Definite_IE", "Healthy", "Highrisk_IE"),
                                          values=c("skyblue","orange","purple"))+
                        stat_compare_means(comparisons=my_comparisons, label.y=c(160,180,170))
boxplot_chao1
```

Figure 3.7c

```{r, warning=FALSE, message=FALSE}
boxplot_invsimp<-ggplot(data = Adiv, aes(x = IE_dx, y = InvSimpson, fill=IE_dx)) + 
                  geom_boxplot(width=0.5)+
                  theme_minimal()+
                  theme(legend.position = "none")+
                  stat_summary(fun.y="mean", geom="point", shape=23, size=3, fill="white")+
                  geom_jitter(shape=1, position=position_jitter(0.1))+
                  ggtitle("InvSimpson Index by Diagnosis") +
  scale_x_discrete(name= "Diagnosis", 
                   limits=c("Definite_IE", "Healthy", "Highrisk_IE"),
                   labels=c("Definite IE", "Healthy", "High risk IE"))+
  scale_fill_manual(limits=c("Definite_IE", "Healthy", "Highrisk_IE"),
                    values=c("skyblue","orange","purple"))+
  stat_compare_means(comparisons=my_comparisons, label.y=c(33, 38, 35))
boxplot_invsimp
```

Figure 3.7d

```{r, warning=FALSE, message=FALSE}
boxplot_SEI<-ggplot(data = Adiv, aes(x = IE_dx, y = Evenness, fill=IE_dx)) + 
  geom_boxplot(width=0.5)+
  theme_minimal()+
  theme(legend.position = "none")+
  stat_summary(fun.y="mean", geom="point", shape=23, size=3, fill="white")+
  geom_jitter(shape=1, position=position_jitter(0.1))+
  ggtitle("Shannon Evenness Index by Diagnosis") +
  scale_x_discrete(name= "Diagnosis", 
                   limits=c("Definite_IE", "Healthy", "Highrisk_IE"),
                   labels=c("Definite IE", "Healthy", "High risk IE"))+
  scale_fill_manual(limits=c("Definite_IE", "Healthy", "Highrisk_IE"),
                    values=c("skyblue","orange","purple"))+
  stat_compare_means(comparisons=my_comparisons, label.y=c(0.84, 0.90, 0.87))
boxplot_SEI
```

**Generate tools for beta diversity visualisation**

```{r, results='hide', warning=FALSE, message=FALSE}
library("phyloseq")
library("ggplot2")
library("dplyr")
library("ggpubr")
library("Matrix")
library("reshape2")
library("vegan")

phylo2=phylo_obj
relab_genera = transform_sample_counts(phylo2, function(x) x / sum(x) * 100) 
head(otu_table(relab_genera)[,1:6])

sample_data(phylo2)$IE_dx<- factor((sample_data(phylo_obj)$IE_dx), levels=c("Definite_IE", "Healthy", "Highrisk_IE"))
abrel_bray <- phyloseq::distance(relab_genera, method = "bray")
abrel_bray <- as.matrix(abrel_bray)
head(abrel_bray)[,1:6] 

sub_dist <- list()
relab_genera = transform_sample_counts(phylo2, function(x) x / sum(x) * 100) ### run again to ensure the loop works
groups_all <- sample_data(relab_genera)$IE_dx

##### below for-loop must all be highlighted to run otherwise df.bray shows error ######
for (group in levels(groups_all)) { 
  row_group <- which(groups_all == group)
  sample_group <- sample_names(relab_genera)[row_group]
  sub_dist[[group]] <- abrel_bray[sample_group, sample_group]
  sub_dist[[group]][!lower.tri(sub_dist[[group]])] <- NA
}

braygroups<- melt(sub_dist)
df.bray <- braygroups[complete.cases(braygroups), ]
df.bray$L1 <- factor(df.bray$L1, levels=names(sub_dist))

head(df.bray)
```

Figure 3.7e Beta diversity box plot

```{r, warning=FALSE, message=FALSE}
df.bray$Diagnosis<-df.bray$L1
betdiv_boxplot1<-ggplot(df.bray, aes(x=Diagnosis, y=value, col=Diagnosis)) +
                    geom_jitter() + 
                    geom_boxplot(alpha=0.6) +  
                    ylab("Bray-Curtis diversity") +
                    xlab("Diagnosis")+
                    ggtitle("Bray-Curtis Diversity in Salivary Microbiome of \nHealthy Samples and Cardiac Patients")+
                    theme_minimal()+
                    theme(plot.title=element_text(hjust=0.5),
                          legend.position="none",
                          axis.text.x=element_text(hjust=0.5,vjust=1,size=10), 
                          axis.text.y=element_text(size=10))+
                    scale_color_manual(limits=c("Definite_IE", "Healthy", "Highrisk_IE"),
                                      values=c("skyblue", "orange", "purple"))+
                    scale_x_discrete(limits=c("Definite_IE", "Healthy", "Highrisk_IE"), 
                                     labels=c("Definite IE", "Healthy",  "High risk IE"))+
                    stat_compare_means(method="wilcox.test", 
                                       comparisons = my_comparisons, 
                                       label.y = c(1.0, 1.10, 1.05))

betdiv_boxplot1
```

Figure 3.7f Beta diversity PCoA plot

```{r, warning=FALSE, message=FALSE}
ord = ordinate(relab_genera, method="PCoA", distance = "bray")
betdiv_pcoa<-plot_ordination(relab_genera, ord, 
                              color = "IE_dx", 
                              shape="IE_dx") + 
                            geom_point(size=3) + 
                            stat_ellipse(aes(group=IE_dx))+
                            theme_minimal()+
                            theme(legend.position = "top",
                                  plot.title=element_text(hjust=0.5))+
                            scale_color_manual(limits=c("Definite_IE", "Healthy", "Highrisk_IE"),
                            values=c("skyblue", "orange", "purple"))+
                            scale_fill_manual(name="Diagnosis")+
                            ggtitle("PCoA Plot of Salivary Microbial Diversity in \nHealthy Samples and Cardiac Patients")
  

betdiv_pcoa
```
Generate heattree package

```{r, results='hide',warning=FALSE, message=FALSE}
library("metacoder")
s=main

names(s)[names(s) == 'X272IE_S111_profile_IEcount.txt'] <- "272IE"
names(s)[names(s) == 'X273IE_S112_profile_IEcount.txt'] <- "273IE"
names(s)[names(s) == 'X290IE_S113_profile_IEcount.txt'] <- "290IE"
names(s)[names(s) == 'X293IE_S13_profile_IEcount.txt'] <- "293IE"
names(s)[names(s) == 'X294IE_S114_profile_IEcount.txt'] <- "294IE"
names(s)[names(s) == 'X295IE_S14_profile_IEcount.txt'] <- "295IE"
names(s)[names(s) == 'X296IE_S115_profile_IEcount.txt'] <- "296IE"
names(s)[names(s) == 'SRS013942_profile_count.txt'] <- "SRS013942"
names(s)[names(s) == 'SRS014468_profile_count.txt'] <- "SRS014468"
names(s)[names(s) == 'SRS014692_profile_count.txt'] <- "SRS014692"
names(s)[names(s) == 'SRS015055_profile_count.txt'] <- "SRS015055"
names(s)[names(s) == 'SRS019120_profile_count.txt'] <- "SRS019120"
names(s)[names(s) == 'SRS065518_profile_count.txt'] <- "SRS065518"
names(s)[names(s) == 'SRS104275_profile_count.txt'] <- "SRS104275"
names(s)[names(s) == 'SRS147126_profile_count.txt'] <- "SRS147126"
s=as.data.frame(s[,-which(names(s) %in% c("SRS015055", "SRS019120"))])

taxmapobj <- parse_tax_data(s,
                            class_cols = "X.clade_name", # the column that contains taxonomic information
                            class_sep = "|", # The character used to separate taxa in the classification
                            class_regex = "^(.+)__(.+)$", # Regex identifying where the data for each taxon is
                            class_key = c(tax_rank = "info", # A key describing each regex capture group
                                          tax_name = "taxon_name"))

print(taxmapobj)


taxmapobj$data$tax_data <- zero_low_counts(taxmapobj, dataset = "tax_data", min_count = 5)

metadata2=Metadata
print(metadata2)
no_reads <- rowSums(taxmapobj$data$tax_data[, metadata2$SampleID]) == 0
sum(no_reads)

taxmapobj <- filter_obs(taxmapobj, target = "tax_data", ! no_reads, drop_taxa = TRUE)
print(taxmapobj)

taxmapobj$data$tax_abund <- calc_taxon_abund(taxmapobj, "tax_data",
                                             cols = metadata2$SampleID)
print(taxmapobj)

taxmapobj$data$tax_occ <- calc_n_samples(taxmapobj, "tax_abund", 
                                         groups = metadata2$IE_dx, 
                                         cols = metadata2$SampleID)

set.seed(1) # This makes the plot appear the same each time it is run 
```

Figure 3.8a

```{r, warning=FALSE, message=FALSE}
heattreeHHS<-heat_tree(taxmapobj, 
                              node_label = taxon_names,
                              node_size = n_obs,
                              node_size_range = c(0.01,0.05),
                              edge_size_range = c(0.005, 0.005),
                              node_color = Healthy,
                              title = "Bacteria Species in Saliva of Healthy Human Samples",
                              title_size = 0.04,
                              node_size_axis_label = "OTU count",
                              node_color_axis_label = "Samples with reads",
                              layout = "davidson-harel", # The primary layout algorithm
                              initial_layout = "reingold-tilford")# The primary layout algorithm
heattreeHHS
```

Figure 3.8b

```{r, warning=FALSE, message=FALSE}
heattreeIE<-heat_tree(taxmapobj, 
                      node_label = taxon_names,
                      node_size = n_obs,
                      node_size_range = c(0.01,0.05),
                      edge_size_range = c(0.005, 0.005),
                      node_color = Definite_IE,
                      title = "Bacteria Species in Saliva of Patients with Definite IE",
                      title_size = 0.04,
                      node_size_axis_label = "OTU count",
                      node_color_axis_label = "Samples with reads",
                      layout = "davidson-harel", # The primary layout algorithm
                      initial_layout = "reingold-tilford")# The primary layout algorithm

heattreeIE
```

Figure 3.8b

```{r, warning=FALSE, message=FALSE}
heattreehighriskIE<-heat_tree(taxmapobj, 
                              node_label = taxon_names,
                              node_size = n_obs,
                              node_size_range = c(0.01,0.05),
                              edge_size_range = c(0.005, 0.005),
                              node_color = Highrisk_IE,
                              title = "Bacteria Species in Saliva of Patients with High Risk IE",
                              title_size = 0.04,
                              node_size_axis_label = "OTU count",
                              node_color_axis_label = "Samples with reads",
                              layout = "davidson-harel", # The primary layout algorithm
                              initial_layout = "reingold-tilford")# The primary layout algorithm
heattreehighriskIE
```

Figure 3.9a 

```{r, warning=FALSE, message=FALSE}
abundance_phylum<-ggplot(data = pd, aes(x = Phylum, y = Abundance)) + 
                    geom_bar(stat = 'identity', aes(fill = IE_dx),width = 0.5) + 
                    theme_minimal()+
                    theme(plot.title=element_text(size=12),
                                legend.position="top",
                                legend.title = element_text(size=11),
                                legend.text=element_text(size=8),
                                legend.key.size=unit(0.5,"cm"),
                                axis.text = element_text(size = 12),
                                axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5)) +
                            scale_fill_manual(name='Diagnosis',limits=c("Definite_IE","Healthy", "Highrisk_IE"), 
                                              labels=c("Definite IE","Healthy", "High risk IE"), values=c("skyblue", "orange", "purple"))+
                    ggtitle("Oral Microbial Abundance by Phylum") +
                    xlab("Phylum")+
                    ylab("Abundance")

abundance_phylum
```

Figure 3.9b

```{r, warning=FALSE, message=FALSE}
pd$IE_dx = factor(pd$IE_dx, levels=c("Healthy", "Definite_IE", "Highrisk_IE"))
abundance_genus<-ggplot(data = pd, aes(x = Genus, y = Abundance)) + 
  geom_bar(stat = 'identity', aes(fill = IE_dx),width = 0.5) + 
  theme_minimal()+ 
  theme(plot.title=element_text(size=20),
        legend.position="top",
        legend.title = element_text(size=11),
        legend.text=element_text(size=8),
        legend.key.size=unit(0.5,"cm"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5)) +
  scale_fill_manual(name='Diagnosis',limits=c("Definite_IE","Healthy", "Highrisk_IE"), 
                    labels=c("Definite IE","Healthy", "High risk IE"), values=c("skyblue", "orange", "purple"))+
  ggtitle("Oral Microbial Abundance by Genus")+
  xlab("Genus")+
  ylab("Abundance")

abundance_genus
```

Figure 3.9c

```{r, warning=FALSE, message=FALSE}
relabundance_diagnosis1<-ggplot(data = pd, aes(x = Phylum, y = Abundance)) + 
  geom_bar(stat = 'identity', position= "fill", aes(fill = IE_dx),width = 0.5) +
  theme_minimal()+
  theme(plot.title=element_text(size=12),
        legend.position="top",
        legend.title = element_text(size=11),
        legend.text=element_text(size=8),
        legend.key.size=unit(0.5,"cm"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5)) +
  labs(x = "Clinical Diagnosis", y = "Relative abundance (%)")+
  scale_fill_manual(name='Diagnosis',limits=c("Definite_IE","Healthy", "Highrisk_IE"), 
                    labels=c("Definite IE","Healthy", "High risk IE"), values=c("skyblue", "orange", "purple"))+
  ggtitle("Relative Abundance by Diagnosis (Phylum)") +
  xlab("Phylum")+
  ylab("Abundance")

relabundance_diagnosis1
```

Figure 3.9d

```{r, warning=FALSE, message=FALSE}
px<-pd
px$Genus[px$Genus==0]=NA
relabundance_diagnosis2<-ggplot(data = pd, aes(x = Genus, y = Abundance, na.rm=T)) + 
  geom_bar(stat = 'identity', position= "fill", aes(fill = IE_dx),width = 0.5) +
  theme_minimal()+
  theme(plot.title=element_text(size=20),
        legend.position="top",
        legend.title = element_text(size=11),
        legend.text=element_text(size=8),
        legend.key.size=unit(0.5,"cm"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5)) +
  labs(x = "Clinical Diagnosis", y = "Relative abundance (%)") +
  scale_fill_manual(name="Diagnosis",limits=c("Definite_IE","Healthy", "Highrisk_IE"), 
                    labels=c("Definite IE","Healthy", "High risk IE"), values=c("skyblue", "orange", "purple"))+
  ggtitle("Relative Abundance by Diagnosis (Genus)") +
  xlab("Genus")+
  ylab("Abundance")

relabundance_diagnosis2
```

Figure 3.10a

```{r, warning=FALSE, message=FALSE}
heatmap_allspecies<-plot_heatmap(phylo_obj, method = "NMDS", distance = "bray",
                          sample.label = "SampleID", taxa.label = "Species", 
                          taxa.order = "Species", na.value="white", low="white", high="red") +
                          theme(legend.position = "top", legend.text =element_text(size=8), legend.key.width = unit(1.5, "cm"))+
                          ggtitle("Species abundance across all samples (NMDS)")+
                          scale_x_discrete(limits=c("272IE","290IE", "294IE", "295IE","SRS014468", "SRS014692", "SRS013942", "SRS147126", "SRS104275", "SRS065518", "273IE", "293IE", "296IE"))
heatmap_allspecies
```

show most abundant OTU

```{r, results='hide', warning=FALSE, message=FALSE}
sample_sums(phylo_obj)
median(sample_sums(phylo_obj))
total = median(sample_sums(phylo_obj))
abundant_OTU2 = phyloseq::filter_taxa(phylo_obj, function(x) sum(x>total*0.20)>0, TRUE)
abundant_OTU2

IE_abundant = subset_samples(abundant_OTU2, IE_dx=="Definite_IE")
HR_abundant = subset_samples(abundant_OTU2, IE_dx=="Highrisk_IE")
HHS_abundant = subset_samples(abundant_OTU2, IE_dx=="Healthy")
```

Figure 3.10b

```{r, warning=FALSE, message=FALSE}
IE_abundant_heatmap2 <- plot_heatmap(IE_abundant, method = "NMDS", 
                                    distance = "euclidean", 
                                    taxa.label = "Species", 
                                    taxa.order = "Species", 
                                    trans=NULL, 
                                    low="white", 
                                    high="skyblue", 
                                    na.value="white", 
                                    sample.label = "SampleID")+
                                    ggtitle("Abundant species in \nDefinite IE samples(NMDS)")+
                                    theme(plot.title=element_text(size=17),
                                      legend.position = "top", 
                                          legend.text =element_text(size=8), 
                                          legend.key.width = unit(1.5, "cm"), 
                                          axis.text.x = element_text(angle = 90, 
                                                                     hjust = 1, 
                                                                     vjust=0.5),
                                          axis.text.y=element_text(size=11))
IE_abundant_heatmap2
```

Figure 3.10c

```{r, warning=FALSE, message=FALSE}
HR_abundant_heatmap2<-plot_heatmap(HR_abundant, method = "NMDS",
                                   distance = "euclidean", 
                                   taxa.label = "Species", 
                                   taxa.order = "Species", 
                                   trans=NULL, 
                                   low="white", 
                                   high="purple", 
                                   na.value="white", 
                                   sample.label = "SampleID")+
                                  ggtitle("Abundant species in \nHigh Risk IE samples(NMDS)")+
  theme(plot.title=element_text(size=17),
        legend.position = "top", 
        legend.text =element_text(size=8), 
        legend.key.width = unit(1.5, "cm"), 
        axis.text.x = element_text(angle = 90, 
                                   hjust = 1, 
                                   vjust=0.5),
        axis.text.y=element_text(size=11))
HR_abundant_heatmap2
```

Figure 3.10d

```{r, warning=FALSE, message=FALSE}
HHS_abundant_heatmap2<-plot_heatmap(HHS_abundant, method = "NMDS",
                                   distance = "euclidean", 
                                   taxa.label = "Species", 
                                   taxa.order = "Species", 
                                   trans=NULL, 
                                   low="white", 
                                   high="orange", 
                                   na.value="white", 
                                   sample.label = "SampleID")+
                                  ggtitle("Abundant species in \nHHS samples(NMDS)")+
  theme(plot.title=element_text(size=17),
        legend.position = "top", 
        legend.text =element_text(size=8), 
        legend.key.width = unit(1.5, "cm"), 
        axis.text.x = element_text(angle = 90, 
                                   hjust = 1, 
                                   vjust=0.5),
        axis.text.y=element_text(size=11))
HHS_abundant_heatmap2
```

Figure 3.12a

```{r, warning=FALSE, message=FALSE}
HHS_netplot<-plot_net(HHS_abundant, 
                     distance = "(A+B-2*J)/(A+B)", 
                     type = "taxa",
                     maxdist = 0.5, 
                     color="Phylum", 
                     point_label="Species", 
                     point_size = 7, 
                     point_alpha=0.5,
                     laymeth="circle", 
                     hjust=0.5,
                     rescale = 0.3)+
                    theme(legend.position = "none")+
                    ggtitle("Network plot for Salivary Microbial Species in HHS Samples")
HHS_netplot
```

Figure 3.12b

```{r, warning=FALSE, message=FALSE}
IE_netplot<-plot_net(IE_abundant, 
                     distance = "(A+B-2*J)/(A+B)", 
                     type = "taxa",
                     maxdist = 0.5, 
                     color="Phylum", 
                     point_label="Species", 
                     point_size = 7, 
                     point_alpha=0.5,
                     laymeth="circle", 
                     hjust=0.5,
                     rescale = 0.3)+
  theme(legend.position = "none")+
              ggtitle("Network plot for Salivary Microbial Species in Definite IE Samples")
IE_netplot
```

Figure 3.12c

```{r, warning=FALSE, message=FALSE}
HR_netplot<-plot_net(HR_abundant, 
                     distance = "(A+B-2*J)/(A+B)", 
                     type = "taxa",
                     maxdist = 0.5, 
                     color="Phylum", 
                     point_label="Species", 
                     point_size = 7, 
                     point_alpha=0.5,
                     laymeth="circle", 
                     hjust=0.5,
                     rescale = 0.3)+
  theme(legend.position = "none")+
        ggtitle("Network plot for Salivary Microbial Species in High risk IE Samples")
HR_netplot
```
